---
title: ''
output: word_document
bibliography: Zotero_references.bib
csl: new-phytologist.csl 

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}

# MAIN MANUSCRIPT DISPLAY ITEMS  (limit 8)
#figures
fig_schematic = 1
fig_NEON_vertical =  2
fig_leaf_T = 3
fig_treerings= 4 
fig_model=5

#tables
table_leaf_traits=1
table_leaf_metabolism=2


#Appendices
appendix_NEON_methods <- "S1"
appendix_tealeaves_methods <- "S2"

appendix_NEON_methods_title<- "Methods for NEON vertical profiles"
appendix_tealeaves_methods_title <- "Methods for leaf energy balance modeling"


#SI table numbers
Tealeaves_parameters = "S1"

#SI table captions
Tealeaves_parameters_legend <- "Table S1. Tealeaves parameters"

#SI figure numbers
SIfig_NEON_all = "S1"

#SI figure captions
SIfig_NEON_all_legend <- "Figure S1. Vertical gradients in micrometeorological conditions for all forested sites in the National Ecological Observatory Network (NEON)"

```


**Title:** Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

**Authors:** 
Nidhi Vinod^1^,
Martijn Slot^2^,
Ian McGregor^3^,
Elsa M. Ordway^4^,
Marielle N. Smith^5^,
Tyeen Taylor,
Lawren Sack,
Kristina J. Anderson-Teixeira^1,2^* 

\newpage

#  Appendix `r appendix_NEON_methods`. `r appendix_NEON_methods_title`

S1 Note:

It is noteworthy that leaf and branch temperatures often differ substantially. For instance, exposed tropical tree bark can be much warmer than leaves [@pauTropicalForestTemperature2018; @stillImagingCanopyTemperature2021; @johnstonWhatLiesVertical2020]. 
Thus, vegetation temperature ($T_{veg}$) measured using sensors that integrate across both leaves and woody vegetation (e.g., infrared sensors for $T_{veg}$ at NEON sites, Fig. `r fig_NEON_vertical`h) does not always equal $T_{leaf}$, particularly in settings where leaf area is low relative to woody vegetation (*i.e.*, low-LAI ecosystems and understories). 
Recent advances in $T_{leaf}$ modeling (see "Scaling across Space and Time") will enhance our understanding of large scale $T_{leaf}$ vertical patterns [@stillImagingCanopyTemperature2021; @macleanMicroclimcMechanisticModel2021]

METHODS ON NEON ANALYSIS GO HERE.

see [Issue 2](https://github.com/EcoClimLab/vertical-thermal-review/issues/2) , [issue 20](https://github.com/EcoClimLab/vertical-thermal-review/issues/20) 

\newpage

#  Appendix `r appendix_tealeaves_methods`. `r appendix_tealeaves_methods_title`

Energy balance for a typical overstory sun leaf and understory shade leaf were modeled in the R package *Tleaves* [@muir_tealeaves_2019], parameterized for *Quercus rubra* L. leaves at Harvard Forest, MA, USA (**GEOGRAPHIC COORDINATES).
Parameters are presented in Table `r Tealeaves_parameters`.

*Describe where you came up with all the parameters:*

- micromet from NEON (explain exactly what you used)
- leaf characteristic dimension, was measured on *Quercus rubra* sun and shade leaf, 
- stomatal conductance measurements were referred from Tleaves typical sun and shade measurements [@muir_tealeaves_2019] 
- @cavender-baresChangesDroughtResponse2000. 

For drought scenario, biophysical variables were parameterized similar to normal scenario using Harvard NEON data. Overstory drought PAR values reflect maximum observed Harvard NEON PAR and understory drought reflect 50% increased PAR of understory normal value. Leaf trait measurements for leaf width is the same as normal scenario, stomatal conductance is kept constant at a minimum value of 0.01 umol/m^2/s/Pa for both positions. In each visual, all variables are constant (Biophysical Constants table) except for the independent variable that represents minimum - maximum range.

\newpage
# `r Tealeaves_parameters_legend `

Make a table from this:
https://github.com/EcoClimLab/vertical-thermal-review/blob/master/NEON_height_profiles/HARV_neon_stats.csv

\newpage

# `r SIfig_NEON_all_legend`
![**Figure `r SIfig_NEON_all`. Vertical gradients in micrometeorological conditions for all forested sites in the National Ecological Observatory Network (NEON)**. Shown are  height profiles in July mean ± 1 standard deviation for maximum photosyntehtically active ratiation (PAR), maximum wind speed, minimum humidity, maximum $T_{air}$, and maximum biological temperature, $T_{bio}$. Sites are grouped into the following categories: (A) (sub)subtropical and warm temperate broadleaf: ..., (B) temperate open/ savanna forests: ..., (C) temperate mesic broadleaf forests: ..., (D) temperate conifer forests: ..., (E) northern and boreal forests: ...  [Issue #35](https://github.com/EcoClimLab/vertical-thermal-review/issues/35). ](NEON_height_profiles/figures/profile_all_groups.png)

\newpage

## References


