---
title: ''
output: word_document
bibliography: Zotero_references.bib
csl: new-phytologist.csl 

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}

#Appendices
appendix_NEON_methods <- "S1"
appendix_tealeaves_methods <- "S2"
appendix_lit_review_methods <- "S3"


appendix_NEON_methods_title<- "Methods for NEON vertical profiles"
appendix_tealeaves_methods_title <- "Methods for leaf energy balance modeling"
appendix_lit_review_methods_title<- "Methods for literature review"



#SI table numbers
SItable_example = "S1"

#SI table captions
SItable_example_legend <- "Table S1. ...."

#SI figure numbers
SIfig_example = "S1"

#SI figure captions
SIfig_example_legend <- "Figure S1. ...."

```


**Title:** Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

**Authors (so far, not necessarily in final order):** 
Nidhi Vinod^1^,
Martijn Slot^2^,
Ian McGregor^3^,
Elsa M. Ordway^4^,
Marielle N. Smith^5^,
Tyeen Taylor,
Lawren Sack,
Kristina J. Anderson-Teixeira^1,2^* 

\newpage

#  Appendix `r appendix_NEON_methods`. `r appendix_NEON_methods_title`

METHODS ON NEON ANALYSIS GO HERE.

see [Issue 2](https://github.com/EcoClimLab/vertical-thermal-review/issues/2) , [issue 20](https://github.com/EcoClimLab/vertical-thermal-review/issues/20) 

\newpage

#  Appendix `r appendix_tealeaves_methods`. `r appendix_tealeaves_methods_title`

METHODS ON TEALEAVES GO HERE.

Energy balance for typical overstory sun leaf and understory shade leaf were parameterized with biophysical variables from Harvard NEON data documented here: (https://github.com/EcoClimLab/vertical-thermal-review/blob/master/NEON_height_profiles/HARV_neon_stats.csv). Leaf trait measurements for effective leaf width was measured on ~Quercus rubra~ sun and shade leaf, stomatal conductance measurements were refered from Tleaves typical sun and shade measurements [@muir_tealeaves_2019] and as well @cavender-baresChangesDroughtResponse2000. Energy balance for biophysical variables in drought scenario were parameterized using Harvard NEON data as well. Leaf trait measurements for effective leaf width is the same as typical scenario, stomatal conductance was referred from @cavender-baresChangesDroughtResponse2000. In each variable visual, all variables are constant except for the independent variable that represents minimum - maximum range.

\newpage

#  Appendix `r appendix_lit_review_methods`. `r appendix_lit_review_methods_title`

METHODS ON LIT REVIEW GO HERE.

\newpage

# Appendix `r appendix_SCBI_treering`. `r appendix_SCBI_treering_title`

\newpage
## References


