---
title: ''
output: word_document
bibliography: Zotero_references.bib
csl: new-phytologist.csl 

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}

#Appendices
appendix_NEON_methods <- "S1"
appendix_tealeaves_methods <- "S2"
appendix_lit_review_methods <- "S3"


appendix_NEON_methods_title<- "Methods for NEON vertical profiles"
appendix_tealeaves_methods_title <- "Methods for leaf energy balance modeling"
appendix_lit_review_methods_title<- "Methods for literature review"



#SI table numbers
SItable_example = "S1"

#SI table captions
SItable_example_legend <- "Table S1. ...."

#SI figure numbers
SIfig_example = "S1"

#SI figure captions
SIfig_example_legend <- "Figure S1. ...."

```


**Title:** Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

**Authors (so far, not necessarily in final order):** 
Nidhi Vinod^1^,
Martijn Slot^2^,
Ian McGregor^3^,
Elsa M. Ordway^4^,
Marielle N. Smith^5^,
Tyeen Taylor,
Lawren Sack,
Kristina J. Anderson-Teixeira^1,2^* 

\newpage

#  Appendix `r appendix_NEON_methods`. `r appendix_NEON_methods_title`

METHODS ON NEON ANALYSIS GO HERE.

see [Issue 2](https://github.com/EcoClimLab/vertical-thermal-review/issues/2) , [issue 20](https://github.com/EcoClimLab/vertical-thermal-review/issues/20) 

\newpage

#  Appendix `r appendix_tealeaves_methods`. `r appendix_tealeaves_methods_title`

METHODS ON TEALEAVES GO HERE.

\newpage

#  Appendix `r appendix_lit_review_methods`. `r appendix_lit_review_methods_title`

METHODS ON LIT REVIEW GO HERE.

\newpage

# Appendix `r appendix_SCBI_treering`. `r appendix_SCBI_treering_title`

\newpage
## References


