---
title: 'Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications'
output:
  word_document:
    toc: true
    toc_depth: 2
bibliography: Zotero_references.bib
csl: new-phytologist.csl 

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}

# MAIN MANUSCRIPT DISPLAY ITEMS  (limit 8)
#figures
fig_schematic = 1
fig_NEON_vertical =  2
fig_leaf_T = 3
fig_treerings= 4 
fig_ecosystem=5
fig_model=6

#tables
table_leaf_traits=1
table_leaf_metabolism=2


# SI
## appendices

appendix_NEON_methods <- "S1"
appendix_tealeaves_methods <- "S2"
appendix_lit_review_methods <- "S3"


appendix_NEON_methods_title<- "Methods S1. Methods for analyzing vertical gradients in the biophysical environment"
appendix_tealeaves_methods_title <- "Methods S2. Methods for leaf energy balance modeling"
appendix_lit_review_methods_title<- "Methods S3. Methods for literature review"



#SI table numbers
NEON_sites = "S1"

#SI table captions
NEON_sites_legend = "Table S1. National Ecological Observatory Network (NEON) sites included in the analysis of vertical gradients of key biophysical characteristics"

#SI figure numbers
SIfig_NEON_all = "S1"

#SI figure captions
SIfig_NEON_all_legend <- "Figure S1. Vertical gradients in micrometeorological conditions for all forested sites in the National Ecological Observatory Network (NEON)"

```


\newpage

##  `r appendix_NEON_methods_title`


Google doc: https://docs.google.com/document/d/1HdPa7X5AMrEwfLjazGMU92uTYMc6pObh/edit#heading=h.gjdgxs 

see [Issue 2](https://github.com/EcoClimLab/vertical-thermal-review/issues/2) , [issue 20](https://github.com/EcoClimLab/vertical-thermal-review/issues/20) 


# Study sites
We analyzed vertical gradients in key biophysical variables from the National Ecological Observatory Network's Airborne Observatory Platform (NEON AOP) and meteorology towers at six focal NEON sites [Figs 2, S1, @schimelNEONHierarchicallyDesigned2007]. The sites are well distributed across the United States and represent key forest types and structures (Table S1).

# Estimating forest and light environments from lidar

*Data selection*

We downloaded classified lidar point clouds as .laz files from the NEON (National Ecological Observatory Network, 2020) data portal. For each site, we analyzed forest structure and light environments using lidar data for the 1 x 1 km plot where the meteorology tower is located. If a large proportion of the plot containing the tower was deforested, we used the closest forested pixel instead. We selected the most recent lidar survey available at each site (all conducted in 2019), which most closely temporally-corresponded to the meteorological time series. All sites were surveyed in growing season months (April-July), except for PUUM (January survey), an evergreen site where total leaf area seasonality is expected to be low.

*Lidar data processing*

Lidar data were processed according to methods detailed in @starkAmazonForestCarbon2012. Outliers on the z plane (height) were removed by applying an interquartile range (IQR) approach. The IQR is calculated as the third quartile (Q3) minus the first quartile (Q1); lidar pulse returns were restricted to > Q1-(1.5xIQR) and < Q3+(1.5xIQR). Next, pulse return counts were quantified, or 'voxelized' within all 2 x 2 x 1 m voxels in a plot.

Digital terrain models (DTMs) were created for each plot using minimum quantile spline regression and interpolation (see Shao et al. 2019) and used to normalize the voxelized lidar data with respect to the ground height. Leaf area density (LAD, $m^2$ $m^{-3}$) was estimated by applying the MacArthur-Horn transformation [@macarthurFoliageProfileVertical1969] to the voxelized pulse returns. We excluded LAD data at the bottom 5 m of the canopy, which are more uncertain due to limited pulse penetration within this region furthest away from the sensor [@starkAmazonForestCarbon2012]; for calibration purposes, we assumed the bottom 5 m to have a leaf area index (LAI) of 0.25 $m^2$ $m^{-2}$ [@shaoHighThroughputAssessment2019].

We scaled LAD values by a factor of 5.5 so that they matched an independent estimate of LAI. Specifically, published estimates of LAI at SERC (average of 5.78 $m^2$ $m^{-2}$) derived from litterfall [@parkerCanopySurfaceStand2004], which were then in good agreement with our lidar-derived estimate (5.75 $m^2$ $m^-2$).
Generating leaf area and light transmission profiles
Mean LAD profiles for each site were calculated as the mean LAD at each height. Mean light transmission profiles (proportion of incident light) were estimated by applying a vertical light reduction model to LAD estimates [@starkAmazonForestCarbon2012]. To estimate the proportion of sun leaves along the vertical profile, we isolated voxels in the upper canopy surface layer (1-2 m from the local canopy surface, likely to containing the majority of sun leaves), calculated the mean LAD profile for this surface layer, and divided it by the total mean LAD profile.

# Analyzing biophysical gradients from micrometeorological data

Micrometeorological data was downloaded for each site from NEON (neonUtilies R package) at 30 minute intervals for 2015-2020. To focus on the middle of the growing season, the data was constrained to be for the month of July each year. Per site, we calculated the mean maximum and minimum values of each variable by day and by sensor height. In Figure S1 we expanded on Figure XX by including 22 sites representing five forest structure types.
 
All data processing, analyses, and graphing were conducted using R (R core team 2020).

*Note on interpreting $T_{veg}$*:

Vegetation temperature ($T_{veg}$) is measured using infrared sensors that integrate across both leaves and woody vegetation. 
Leaf and branch temperatures often differ substantially; for instance, exposed tropical tree bark can be much warmer than leaves [@pauTropicalForestTemperature2018; @stillImagingCanopyTemperature2021; @johnstonWhatLiesVertical2020].
Thus, vegetation temperature ($T_{veg}$) does not always equal or approximate $T_{leaf}$, particularly in settings where leaf area is low relative to woody vegetation (i.e., low-LAI ecosystems and understories).


\newpage

##  `r appendix_tealeaves_methods_title`

Energy balance for a typical overstory sun leaf and understory shade leaf were modeled in the R package *tealeaves* [@muir_tealeaves_2019], parameterized for *Quercus rubra* L. leaves at Harvard Forest, MA, USA (42.5369, -72.17266).

Micrometeorological data from the NEON tower at Harvard Forest (Fig. `r fig_NEON_vertical`, Supporting Information Methods `r appendix_NEON_methods`) was used to parameterize biophysical constants, using the mean of maximum PAR, wind speed, and $T_{air}$ for growing season months at vertical positions 60m (overstory) and 10m (understory). 
Leaf size parameters were selected for *Quercus rubra* sun and shade leaf based on personal observations (by Vinod) and verified as reasonable by @zwienieckiHydraulicLimitationsImposed2004. 

For the humid scenario, we used mean maximum RH from the NEON tower at Harvard Forest.
Stomatal conductance measurements were referred from *Tleaves* typical sun and shade measurements [@muir_tealeaves_2019] and @cavender-baresChangesDroughtResponse2000. 

For the drought scenario, we reduced RH to 50% of the mean maximum from the NEON tower at Harvard Forest.
For drought PAR values, for the overstory we used the maximum PAR value observed at 60m height on the NEON tower at Harvard Forest, and for the understory we assumed 50% increased PAR relative to the mean daily maximum understory value.
We reduced stomatal conductance to a minimum value of 0.01 $\mu mol/m^2/s/Pa$ for both overstory and understory. 

In each visual, all biophysical variables are constant except for the independent variable that represents a minimum - maximum range. 

\newpage
# `r appendix_lit_review_methods_title`

## Identification of relevant studies

Relevant studies were searched in the databases: ISI Web of Science, Smithsonian online library and Google Scholar, using the following key terms: (leaf traits OR foliar traits) AND (inter-canopy OR intra-canopy OR canopy height) AND (e.g. chlorophyll OR e.g. LMA OR stomatal conductance); (leaf temperature and metabolism OR leaf thermal sensitivity  OR leaf thermal tolerance OR leaf traits OR foliar traits) AND (within-canopy OR intra-canopy OR sun shade OR canopy height OR canopy gradient OR canopy profile OR canopy position) AND (temperate forests OR boreal forest OR conifer OR savanna OR tropical); (leaf* temperature* and metabolism OR leaf thermal* sensitivity  OR leaf thermal tolerance OR leaf* traits OR foliar* traits) AND (within-canopy OR intra-canopy OR sun shade OR canopy* height OR canopy gradient OR canopy profile) AND (temperate forests OR boreal forest OR conifer OR savanna OR tropical).

in Google Scholar, 600 articles were screened, using the key terms (leaf temperature and metabolism OR leaf thermal sensitivity  OR leaf thermal tolerance OR leaf traits OR foliar traits) AND (within-canopy OR intra-canopy OR sun shade OR canopy height OR canopy gradient OR canopy profile OR canopy position) AND (temperate forests OR boreal forest OR conifer OR savanna OR tropical). After screening the title and reading the abstract, 185 articles were most relevant. Herbaceous plant studies and seedling studies were excluded. These articles were added into Zotero folders: Leaf traits;  Leaf Metabolism and Processes, accordingly, for further careful reading.

Similar search was conducted with the Smithsonian online library with the key terms (leaf traits OR foliar traits) AND (inter-canopy OR intra-canopy OR canopy height) AND (e.g. chlorophyll OR e.g. LMA OR stomatal conductance) provided most relevant results among the set of keywords mentioned above, with the field refine function that included botany, ecology, biology, environmental sciences, and forestry. Out of 150 relevant articles screened, 26 were most relevant after screening the title and reading the abstract, out of which 22 were already acquired through Google scholar search, 4 new articles were added to the folder. 

For ISI Web of Science, search the key terms (leaf* temperature* and metabolism OR leaf thermal* sensitivity  OR leaf thermal tolerance OR leaf* traits OR foliar* traits) AND (within-canopy OR intra-canopy OR sun shade OR canopy* height OR canopy gradient OR canopy profile) AND (temperate forests OR boreal forest OR conifer OR savanna OR tropical) yielded 410 relevant results, 37 were most relevant after screening the title and reading the abstract, out of which 24 were already acquired through the above process, 13 new articles were added to Zotero folders.

Through the above process, 202 articles were acquired into Zotero folders. The articles were tagged after careful reading of each as added to the table, irrelevant to the table, but relevant to the reviewed (with a note on the reason), irrelevant to the table and the review (with a note on the reason for exclusion). Articles that were not yet carefully read were tagged yet to read, which will be tagged as one of the above after reading. Articles shared by co-authors and references mentioned in other studies collectively were ~>32 studies. So far 70 relevant articles are added into the tables

## Criteria for inclusion 

**this is outlined [here](https://github.com/EcoClimLab/vertical-thermal-review/blob/master/Lit_Review/Methods%20for%20Systematic%20review%20for%20the%20tables.docx). Section needs to be written.**

The studies included in the review were based on a global geographic scope. Ecosystem types included any studies with trees for mostly forests, but savannas were also considered. Information for leaf traits and metabolism included studies with independent variables: sun/shade vs height (where height is generally also sun/shade), and dependent variables of interest were based on variables known to influence leaf temperature ie. anatomical, biochemical, structural foliar traits, including gas exchange, metabolism and thermal sensitivity variables. 

\newpage
 # `r NEON_sites_legend`

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 9)
NEON_table <- read.csv("NEON_height_profiles/NEON_sites_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
NEON_table <- NEON_table[c(2:4, 6:7,9)] #select and order columns for inclusion
ft <- flextable(NEON_table)
ft <- autofit(ft)
ft <- width(ft, 1, width = .75)
ft <- width(ft, 2, width = 1.75)
ft <- width(ft, 3, width = .5)
ft <- width(ft, 4, width = .75)
ft <- width(ft, 5, width = .8)
ft <- width(ft, 6, width = 2)
ft
```
*Refers to forest type around the tower on which the vertical profile of micrometeorological instruments is mounted, as provided by NEON (Hongyan Luo, personal communication; [site descriptions on NEON website](https://www.neonscience.org/field-sites)). 
 
\newpage

# `r SIfig_NEON_all_legend`
![**Figure `r SIfig_NEON_all`. Vertical gradients in micrometeorological conditions for all forested sites in the National Ecological Observatory Network (NEON)**. Shown are  height profiles in July mean ± 1 standard deviation for maximum photosyntehtically active ratiation (PAR), maximum wind speed, minimum humidity, maximum $T_{air}$, and maximum biological temperature, $T_{bio}$. Site information is given in Supporting Information Table `r NEON_sites`, and analysis details in Supporting Information Methods `r appendix_NEON_methods`.](NEON_height_profiles/figures/profile_all_groups.png)

\newpage
## References


